### 2022-01-05 paper/figures/masktree.tex
update bounding

\documentclass[crop,tikz]{standalone}

\begin{document}
\begin{tikzpicture}[scale=1.5]
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (0.0, 0.0) (1) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (-0.8, -0.6) (2) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (0.8, -0.6) (3) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (-1.3, -1.2) (4) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (-0.30000000000000004, -1.2) (5) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (1.3, -1.2) (6) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (0.30000000000000004, -1.2) (7) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (-1.6, -1.8) (8) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (-1.0, -1.8) (9) {};
\draw[latex-, line width=1pt] (1) -- (2);
\draw[latex-, line width=1pt] (1) -- (3);
\draw[latex-, line width=1pt] (2) -- (4);
\draw[latex-, line width=1pt] (2) -- (5);
\draw[latex-, line width=1pt] (3) -- (6);
\draw[latex-, line width=1pt] (3) -- (7);
\draw[latex-, line width=1pt] (4) -- (8);
\draw[latex-, line width=1pt] (4) -- (9);
\node[] at (0.0, 0.3) {$\alpha(G)$};
\node[] at (-1.6, -1.2) {$A$};
\node[] at (-0.30000000000000004, -1.5) {$B$};
\node[] at (-0.8, -0.3) {$C$};
\node[] at (-0.8, -0.8999999999999999) {$*$};
\node[] at (0.0, -2.2) {\large (a)};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (3.6, 0.0) (10) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (2.8, -0.6) (11) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (4.4, -0.6) (12) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (2.3, -1.2) (13) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (3.3, -1.2) (14) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (4.9, -1.2) (15) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (3.9000000000000004, -1.2) (16) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (2.0, -1.8) (17) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (2.6, -1.8) (18) {};
\draw[-latex, draw=red, line width=1pt] (10) -- (11);
\draw[-latex, draw=red, line width=1pt] (10) -- (12);
\draw[-latex, draw=red, line width=1pt] (11) -- (13);
\draw[-latex, draw=red, line width=1pt] (11) -- (14);
\draw[-latex, draw=red, line width=1pt] (12) -- (15);
\draw[-latex, draw=red, line width=1pt] (12) -- (16);
\draw[-latex, draw=red, line width=1pt] (13) -- (17);
\draw[-latex, draw=red, line width=1pt] (13) -- (18);
\node[] at (3.6, 0.3) {$\overline{\alpha(G)}$};
\node[] at (1.9999999999999998, -1.2) {$\overline{A}$};
\node[] at (3.3, -1.5) {$\overline{B}$};
\node[] at (2.8, -0.3) {$\overline{C}$};
\node[] at (2.8, -1.0) {\small Eq. 6.9};
\node[] at (3.6, -2.2) {\large (b)};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (7.2, 0.0) (19) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (6.4, -0.6) (20) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (8.0, -0.6) (21) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (5.9, -1.2) (22) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (6.9, -1.2) (23) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (8.5, -1.2) (24) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (7.5, -1.2) (25) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (5.6, -1.8) (26) {};
\node[circle, inner sep=0cm, draw=black, line width=0.03, fill=none, minimum size=0.3cm] at (6.2, -1.8) (27) {};
\draw[latex-, draw=black, line width=1pt] (19) -- (20);
\draw[latex-, draw=black, line width=1pt] (19) -- (21);
\draw[latex-, draw=black, line width=1pt] (20) -- (22);
\draw[latex-, draw=black, line width=1pt] (20) -- (23);
\draw[latex-, draw=black, line width=1pt] (21) -- (24);
\draw[latex-, draw=black, line width=1pt] (21) -- (25);
\draw[latex-, draw=black, line width=1pt] (22) -- (26);
\draw[latex-, draw=black, line width=1pt] (22) -- (27);
\node[] at (7.2, 0.3) {$s_{\alpha(G)}\infty^{\alpha(G)}$};
\node[] at (5.45, -1.2) {$A'\circ\overline{A}$};
\node[] at (6.9, -1.5) {$B'\circ\overline{B}$};
\node[] at (6.4, -0.3) {$C'\circ\overline{C}$};
\node[] at (6.4, -0.8999999999999999) {$*$};
\node[] at (7.2, -2.2) {\large (c)};
\end{tikzpicture}
\end{document}